---
title             : "Complementing theory on adolescent emotion regulation using data-driven insights from machine learning"
shorttitle        : "Title"

author:
  - name: "Caspar J. Van Lissa"
    affiliation: "1,2"
    corresponding: yes
    address: "Padualaan 14, 3584CH Utrecht, The Netherlands"
    email: "c.j.vanlissa@uu.nl"
  - name: "Lukas Beinhauer"
    affiliation: "3"
  - name: "Susan Branje"
    affiliation: "4"
  - name: "Wim H. J. Meeus"
    affiliation: "4"

affiliation:
  - id: "1"
    institution: "Utrecht University, department of Methodology & Statistics"
  - id: "2"
    institution: "Open Science Community Utrecht"
  - id: "3"
    institution: "Helmut-Schmidt-Universität, Department of Methodology and Statistics for Psychology"
  - id: "4"
    institution: "Utrecht University, department of Youth and Family"

authornote: |
  Data of the RADAR (Research on Adolescent Development And Relationships) study were used (https://doi.org/10.17026/dans-zrb-v5wp). RADAR has been financially supported by main grants from the Netherlands Organisation for Scientific Research (GB-MAGW 480-03-005, GB-MAGW 480-08-006), Stichting Achmea Slachtoffer en Samenleving (SASS), the Netherlands Organisation for Scientific Research to the Consortium Individual Development (CID; 024.001.003), a grant of the European Research Council (ERC-2017-CoG - 773023 INTRANSITION), and various other grants from the Netherlands Organisation for Scientific Research, VU University Amsterdam, and Utrecht University. The funders had no role in the study design, data collection and analysis, decision to publish, or preparation of the manuscript. If the regulations of the journal allow it, the participating children, parents and schools are to be thanked as well.

abstract: |
  https://www.srcd.org/news/child-development-special-section-formalizing-theories-child-development
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  
  One or two sentences to put the results into a more **general context**.
  
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib", "references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
knit              : worcs::cite_all
---

```{r setup, include = FALSE}
library("papaja")
library("worcs")
# We recommend that you prepare your raw data for analysis in 'prepare_data.R',
# and end that file with either open_data(yourdata), or closed_data(yourdata).
# Then, uncomment the line below to load the original or synthetic data
# (whichever is available), to allow anyone to reproduce your code:
# load_data()
```

<!-- TS: Emotion regulation development is a critical challenge in adolescence. -->
<!-- TS: Emotion dysregulation has severe consequences for individual adolescents, and society at large. -->
<!-- TS: Therefore, substantial research has addressed this topic. -->
<!-- TS: Most preceding research has been limited in scope, focusing on only a few theoretically-driven predictors. -->
<!-- TS: Machine learning makes it possible to cast a wider net, and explore all (available) potentially relevant predictors. -->
<!-- TS: Confirmatory versus exploratory research: Define the distinction. -->
<!-- TS: Some longitudinal research falls between these categories. For example, complex longitudinal studies, where prior research is sufficient to lead us to hypothesize *some* associations between two constructs, but not sufficiently specific to let us hypothesize which parameters will be significant. -->
<!-- TS: Instead, us -->

A key challenge in adolescence is developing mature emotion regulation abilities [@zimmermannEmotionRegulationEarly2014]. 
Adolescents experience biological, cognitive, and social changes that create an opportunity for positive emotion regulation development [@steinbergAgeOpportunityLessons2014].
Still, as many as 20% of adolescents instead experience severe emotion regulation difficulties [@leeAdolescentMentalHealth2014],
with potentially lifelong implications for mental health and well-being [@aldaoEmotionregulationStrategiesPsychopathology2010].
These difficulties in emotion regulation can cause substantial distress to the individual,
and incur a sizable societal cost.
A comprehensive understanding of the risk factors that predispose some adolescents to emotion dysregulation could aid in the development of effective screening and interventions.
However, despite an abundance of empirical research,
many studies are narrow in scope, considering only a few predictors,
the literature is fragmented across subdisciplines, and lacks a unifying theoretical framework
<!-- that is both specific to adolescence and comprehensive in scope  -->
[@vanlissaMappingPhenomenaRelevant2021].

The present study seeks to overcome these limitations
by using data-driven insights to complement and reflect on existing theory.
To this end, a broad range of potential risk factors were analyzed using machine learning,
and ranked in terms of predictive importance.
<!-- These predictors may include both risk factors that render some youth susceptible to emotion regulation difficulties, and correlates that can be used to identify at-risk youth. -->
Specifically,
the SEM-forests algorithm was used to
identify the most important early indicators of developmental trajectories
of emotion regulation across adolescence.
The resulting insights indicate which factors are most predictive of adolescents' diverging destinies in terms of emotion regulation development.
Armed with this knowledge, future research and interventions can focus on the most important predictors.
Moreover, these insights can guide the formation of overarching theory that adequately represents essential phenomena.
<!-- complement existing theory on adolescent emotion regulation using data-driven insights from machine learning. -->

## 

## Emotion dysregulation in adolescence

Emotion regulation,
defined as the ability to manage emotional experience and expression,
is critical for mental health4–6 and social competence2 throughout life. There is mounting evidence, however, that adolescents do not uniformly become better at regulating emotions1. 
According to Crone and Dahl's neurodevelopmental model1,
staggered development of emotional and regulatory brain circuits gives rise to a "maturity gap" in mid-adolescence:
A period wherein the emotional salience of social stimuli is intensified, while the ability to regulate these emotions lags behind8,21,22.
This imbalance motivates most adolescents to engage in adaptive exploration:
Seeking out experiences required to develop mature socio-emotional competences.
At the same time, this imbalance renders some adolescents vulnerable to emotion dysregulation1,10,
<!-- exposes youth to a risk of emotion dysregulation, -->
for example, because they are not yet well-equipped to regulate the pursuit of unhealthy incentives,
or persevere in pursuing adaptive goals when facing obstacles.
Thus, some become trapped in a negative spiral characterized by more negative and volatile emotions2,9,11,14,
which has been linked to the emergence of several mental disorders5,10.
<!-- This imbalance motivates most youngsters to seek out new challenges in life and love, -->
<!-- but renders some of them vulnerable to emotion dysregulation1,10. Unfortunately, there is little theory about which risk factors and environmental hazards render adolescents susceptible to emotional difficulties11. -->
<!-- Acquiring adaptive emotion regulation skills in adolescence plays a key role in this regard7,20, because t -->
 
Empirical support for this notion of diverging destinies11 is evident from studies that examine developmental trajectories of emotion regulation in adolescence: 
Such studies consistently reveal large between-individual heterogeneity 2,7,9.
An important next step is to identify which risk factors render some adolescents susceptible to emotion dysregulation.
This involves three important challenges:
The first is that the number of potentially relevant risk factors is large,
the second is that there is a paucity of relevant theory to narrow down this number,
and the third is that methods commonly used in the field are not well suited for large numbers of predictors.
We argue that machine learning methods that perform variable selection overcome these challenges.

<!-- This goal fits within the paradigm shift that has gripped the field of pedagogy, towards “person-centered” research, which focuses on individual developmental trajectories and within-family predictive processes, rather than average patterns across samples16,23. Seminal studies have applied “clustering” methods to identify youth with similar developmental trajectories14,24–26. Adolescents may follow similar trajectories for different reasons, however, and person-centered methods are too complex to accommodate more than a handful of theoretically-derived predictors. Unfortunately, relevant theory on adolescent emotion regulation development is limited27–29. The field is therefore at an impasse: We know that some adolescents are more susceptible to emotion dysregulation than others, but lack tools and theory to identify important predictors of individual development30. -->

## Potential predictors

<!-- Maybe review theory briefly? -->
A recent literature review set out to identify relevant phenomena in relation to adolescents' emotion regulation [@vanlissaMappingPhenomenaRelevant2021].
It used a new technique, text mining systematic review,
to identify relevant phenomena based on frequently occurring terms in the author keywords and abstracts of 6305 publications.
Whereas narrative reviews are typically restricted to small corpora and structured around well-established ideas [@littellEvidencebasedBiasedQuality2008],
text mining systematic reviews are more comprehensive and derive insights from the literature through a relatively transparent and objective procedure - at the expense of reflection and meaning-making.

The findings of that systematic review indicated that internalizing and externalizing problems and mental health-related outcomes were most frequently studied in conjunction with emotion regulation. 
Other relevant phenomena reflected themes discussed in prior narrative reviews,
including cognitive development, parenting, and external stressors [cf. @coe-odessEmergentEmotionsAdolescence2019].
Other relevant phenomena were underrepresented in prior theory and reviews,
such as substance use, structural disadvantage,
identity, morality, sex and sexual development.
This relatively comprehensive overview offers a starting point for empirical investigations of predictors of adolescents' emotion regulation development.
<!-- To this end, the present study sought to identify operationalizations of these relevant phenomena in a large-scale panel study. -->

## Complementing theory using machine learning

As mentioned before, the field of adolescent emotion regulation is prolific, but lacks overarching theory.
For guidance on how to remedy this shortcoming, we turn to philosophy of science.
De Groot’s "empirical cycle" constitutes a model of the production of knowledge through scientific research [@degrootMethodologieGrondslagenVan1961].
It consists of five phases: *observation* of phenomena to be explained; *induction* of hypotheses; *deduction* or derivation of testable predictions; *testing* these predictions in new data; and finally *evaluation*: reflecting back on the theories and hypotheses that gave rise to the tested predictions.
The model is cyclical because *evaluation* offers new interpretative data for the researcher to observe - such as rejected hypotheses and serendipitous findings.

<!-- The empirical cycle is merely one model, -->
<!-- and there is a pluriformity of other epistemological approaches to knowledge production. -->
<!-- Nonetheless, u -->
Unsubstantiated or undisclosed deviations from the empirical cycle may signal practices that limit scientific progress.
Consider three examples:
First, conducting many analyses and then writing a paper that presents significant findings as tests of a-priori hypotheses, or "hypothesizing after the results are known", constitutes a questionable research practice that leads to false-positive results [@wichertsDegreesFreedomPlanning2016].
Second, failure to engage in evaluation and reflect back on theory in light of empirical findings leads to flexible, "invulnerable" theories that can accommodate any result [@szollosiArrestedTheoryDevelopment2021].
Third, and particularly relevant to the present study,
is failure to complete the cycle.
In a productive field of study, the empirical cycle is regularly completed;
evaluation of theories in the light of empirical findings allows scholars to hone those theories,
which in turn yield increasingly precise predictions.
We argue that, the paucity of overarching theory in our field can be remedied by completing the empirical cycle.

The phases of the empirical cycle map onto two modes of conducting research:
Confirmatory research, which involves the phases of deduction, testing, and evaluation, and exploratory research, which involves the phases of observation and induction  [@wagenmakersCreativityVerificationCyclePsychological2018].
Although developmental psychology is dominated by quantitative positivism,
relatively few studies are explicitly exploratory [@demuthQualitativeMethodologyDevelopmental2015].


and devoting greater attention to exploratory research and theory formation.



<!-- In this model, the hypothesis plays an important role. -->
<!-- It is a testable prediction that can be supported or falsified; depending on the results, we reflect on old knowledge – which includes theory – and make necessary adjustments that lead to new hypotheses. -->
<!-- A hypothesis can thus be a specific prediction that delivers new knowledge when it is false; -->
<!-- this drives a scientific field forward. -->


Prior theory and empirical research thus suggests many phenomena potentially related to heterogeneity in adolescents' emotion regulation development.
What is missing is an understanding of the relative utility of each of these phenomena in predicting developmental outcomes.
Individual studies do not suffice for this purpose as
they tend to sample from different populations,
use different research designs,
different control variables and statistical models.
Relatedly, many studies tend to examine only a handful of predictors,
often guided by a theory with narrow scope - 
, it is hard to tell how important the variables are, relative to other variables considered by other theories (or variables not considered at all). Machine learning is a more holistic methodology, as it can assess and compare a large number of potential predictors simultaneously, including theoretically relevant ones, and identify which predictors ultimately explain the most variance in the outcome measure of interest
  
  
  Machine learning is a more holistic methodology, as it can assess and compare a large number of potential predictors simultaneously, including theoretically relevant ones, and identify which predictors ultimately explain the most variance in the outcome measure of interest.
  
  
Theory cons methods for theory construction state that the first step in this endeavor is to identify well-established relevant *empirical phenomena*,
defined as stable and general features of the world [@borsboomTheoryConstructionMethodology2020].
The current study used a text mining systematic review to provide a comprehensive map of the empirical phenomena relevant to adolescent emotion regulation.

Currently, few instruments exist for the identification of relevant phenomena; instead, theorists typically rely on expert opinion [@borsboomTheoryConstructionMethodology2020].
This paper argues that systematic reviews are a suitable and relatively more objective method for phenomena detection.
Traditional narrative reviews are limited, however, by small convenience samples,
confirmation bias,
and an undue emphasis on positive results [@littellEvidencebasedBiasedQuality2008].
These limitations are overcome by a novel method:
the text mining systematic review (TMSR).
Relative to narrative reviews, text mining can cover arbitrarily greater corpora,
and gleans insights from the literature by means of a relatively more transparent, objective, and reproducible process.
<!--Where sentient readers tend to structure their reading of the literature around established ideas [@littellEvidencebasedBiasedQuality2008],
the inherent "fairness" of text mining gives emerging themes a chance to come to the fore.-->
This approach assumes the frequency with which phenomena recur in the literature to be indicative of their relevance.
If it is additionally assumed that the frequency with which phenomena are studied together is indicative of the extent to which they are thought to be associated,
then the co-occurrence of phenomena within publications can be interpreted as a rudimentary nomological network: a mapping of relationships between theoretically relevant phenomena [see @alaviAligningTheoryMethodology2018].
Such a network can serve as a starting point for the formulation of a proto-theory,
which would additionally involve abductive reasoning: specifying putative mechanisms to explain associations between the phenomena [@borsboomTheoryConstructionMethodology2020].
Text mining systematic reviews constitute an inductive and exploratory approach:
Their aim is phenomena detection based on generalization from observations (data) without appeal to theory [@haigExploratoryFactorAnalysis2005].
By identifying and mapping the phenomena relevant to adolescent emotion regulation,
such an inductive approach lays the groundwork for future theory development.

<!--
stifterEmotionRegulation2019
* call for studies incorporating predictors of ER across multiple levels of analysis
* different conceptual frameworks

riedigerEmotionRegulationAdolescence2014
* difficulty in deriving a cohesive picture about emotion regulation in adolescence
* different conceptual frameworks

* three themes (neurocognitive development, socialization, interplay between internal and external factors) currently represent relatively independent bodies of research
* A more explicit integration of these various research perspectives will help us to arrive at a more integrated and interpretable picture of the development of emorion regulation during adolescence. 

@bussTheoriesEmotionalDevelopment2019

* emotional development theory is still in its own state of development. 
* empirical evidence can contribute to theory development.

@bariolaChildAdolescentEmotion2011
* Further research apply reﬁned theoretical conceptualizations of ER.


Where sentient readers tend to structure their reading of the literature around established ideas [@littellEvidencebasedBiasedQuality2008],
This potentially reveals phenomena and associations missed by narrative reviewers,
thereby laying the groundwork for future empirical and theoretical work.


we set out to map these relevant empirical phenomena using a systematic review.


generate a proto-theoretical nomological network of the relevant empirical phenomena.
We discuss how this nomological network relates to existing relevant theory, and conclude
with recommendations for future theory-generating efforts.
There have been calls for the  consolidation of these disparate insights into overarching theory would be an important step forward.
An important first step 
and intrinsic limitations of narrative reviews prohibit comprehensive coverage of the vast but diffuse literature .-->
<!--
Recent years have seen the publication of several excellent systematic reviews,
but these have all been narrative reviews.-->

 <!-- CJ To fill out gaps, following up on promising constructs and unify heretofore disparate regions of the lit -->
<!-- SH: this implies that the main contribution of the paper is predominantly methodological is that the case? Is there some primary thesis/argument that will be advanced? It is really important to take a top-down approach to that messaging, beginning here in the first paragraph-->
<!-- Ergens anders: Similarly, although
there are several *relevant* theories, there is no specific theoretical
framework to guide research on adolescents' emotion regulation. Although theory
development is beyond the scope of this paper, we set out to generate a
proto-theoretical nomological network of the potential risk factors,
manifestations, and outcomes of emotion regulation in adolescence. We discuss
how this nomological network relates to existing relevant theory, and conclude
with recommendations for future theory-generating efforts.-->
<!-- At the same time, the staggered development of motivational-emotional and regulatory brain circuits gives rise
to a maturity gap [@croneUnderstandingAdolescencePeriod2012]. 
This leads youngsters to pursue new experiences in life and love,
without being fully prepared to cope with the emotional outcomes. --> 

<!--The literature thus paints a picture of
adolescence as a chrysalis for emotional development: Children enter this stage
with emotion regulation skills adapted to the challenges of childhood. During
adolescence, emotional systems are rearranged substantially.-->

<!--
* intense negative emotions in daily life (Silk, Steinberg, & Morris, 2003) 
* unstable peer or romantic relationships (Furman & Collins, 2009)
* decrease in perceived support from parents (Furman & Buhrmester, 1992).
* early adolescence is characterized by a higher rate of conflicts with parents (Laursen, Coy, & Collins, 1998)
* higher variability of negative emotions compared to late adolescence (Larson, Moneta, Richards, & Wilson, 2002).
* relative to other developmental stages, adolescence involves a particularly large number of transitions, novel situations, and new stimuli, physical, cognitive, emotional, and social development—particularly in early and mid-adolescence (Seiffge-Krenke, 2000)
-->


## Knowledge gap

Substantial empirical research, but:

* Different (sub)disciplines have approached the topic in disparate ways [@riedigerEmotionRegulationAdolescence2014]
* No consistent terminology [@bariolaChildAdolescentEmotion2011]
* No conceptual frameworks [@stifterEmotionRegulation2019]
* No overarching theoretical framework [@bussTheoriesEmotionalDevelopment2019]

Knowledge must be consolidated into overarching theory

## Relevant theory

* Bioecological model [@bronfenbrennerBioecologicalModelHuman2007]
* Transactional model [@sameroffUnifiedTheoryDevelopment2010]
* Hall's notion of "storm and stress"
* Theory of normative emotional development [@sroufeEmotionalDevelopmentOrganization1995]
* Tripartite model [@morrisRoleFamilyContext2007]
* Internalization model [@holodynskiDevelopmentEmotionsEmotion2006]
* Polyvagal theory [@porgesOrientingDefensiveWorld1995].
* Model of social-affective engagement and goal flexibility [@croneUnderstandingAdolescencePeriod2012]
* Process model of emotion regulation [@grossHandbookEmotionRegulation2013]
* Social Information Processing Theory [@lemeriseIntegratedModelEmotion2000]

## Limitations of relevant theory

* Few explicitly address adolescence
* Few comprehensively address predictors of emotion regulation
* None directly guide contemporary research 
* Theories vary widely in scope: Some are broad and non-specific; others describe a specific phenomenon in detail, but lack a broader perspective.
    + Broad theories can frame any research, specific theories generate hypotheses.
    + It would be beneficial to bridge these levels of analysis.

## Towards integrative theory

First step: identifying relevant phenomena [@borsboomTheoryConstructionMethodology2020]

We conducted a text mining systematic review (TMSR) [@vanlissaMappingPhenomenaRelevant2021]

  * Narrative reviews: small samples, confirmation bias, emphasize positive results [@littellEvidencebasedBiasedQuality2008]
  * TMSR: Unlimited sample size, transparent, objective, reproducible

6653 papers on Addresses **emotion regulation** in population overlapping with adolescence [10-24]

<doi.org/10.1007/s40894-021-00160-7>

## Co-occurrence graph

Author keywords (a) and abstracts (b)

## Next step


* Predictors of latent growth model?
  + Only linear differences, few predictors (power)
* Multigroup latent growth model?
  + Non-linear differences, but only one moderator
* Latent class growth model with auxilliary variables?
  + Groups with regard to trajectory, not predictors

## Solution: Machine learning

* Allows non-linear differences between trajectories
* Performs variable selection
  + Cast wide net among potentially relevant predictors
* Checks & balances ensure generalizability
* Exploratory: No hypothesis

## Method illustrated

## Method

SEM-forest [@brandmaierTheoryguidedExplorationStructural2016]

1. Bootstrap sample
2. On each sample, estimate a SEM-model
3. Consider $k$ candidate predictors
4. Identify predictor and value that maximizes $LR$ of post-split multi-group model
5. Average predictions across bootstrap samples

We used 1000 bootstrap samples, $k = 9$

<!-- TS: Confirmatory versus exploratory research: Define the distinction. -->
<!-- TS: Some longitudinal research falls between these categories. For example, complex longitudinal studies, where prior research is sufficient to lead us to hypothesize *some* associations between two constructs, but not sufficiently specific to let us hypothesize which parameters will be significant. -->

<!-- ## Results -->

<!-- * Both analyses reflected some constructs from theoretical literature -->
<!--     + neurodevelopment, socialization -->
<!-- * Undertheorized themes -->
<!--     + Developmental disorders -->
<!--     + Physical health (sic) -->
<!--     + External stressors -->
<!--     + Structural disadvantage -->
<!--     + Addictive behavior -->
<!--     + Identity and moral development -->
<!--     + Sexual development -->

<!-- * This offers guidance for phenomena to consider as relevant predictors -->


## Present study

RQ1: What are the most important predictors of adolescents' trajectories of emotion regulation development?

RQ2: What is the nature of the association of the predictors with the trajectories?

**Participants**

* RADAR data (https://www.uu.nl/en/research/radar)
* 497 Dutch adolescents (283 boys; age at T1: M =13.03, SD = 0.46)
* Collected between 2006-2011
* Most families were  medium- to high-SES (10% low-SES)

## Candidate predictors

Predictors: 87

Categories: 

* Demographics (e.g., SES)
* Biological (age, sex, pubertal development)
* Individual differences (big5, BIS/BAS, empathy)
* Risk behavior (in/externalizing symptoms, substance use, delinquency, school performance, victimization)
* Relationship quality (parents/peers)
* Parenting (perceived and parent-reported)
* Conflict resolution styles

## Dependent variable

Difficulties in emotion regulation [@gratzMultidimensionalAssessmentEmotion2004]

* 24 items (omitted 'emotional awareness' subscale)
* E.g.: “When I’m upset, I have difficulty thinking about anything else”
* 1 (“Almost never”) to 5 (“Almost always”)
* Excellent reliability, $(.94 > \alpha > .95)$.
* Metric invariance across waves
* Quadratic growth model had good fit
  + CFI = .996, TLI = .993, RMSEA = .047, SRMR = 0.020

## Open science

Workflow for Open Reproducible Code in Science (WORCS) used to make analyses reproducible [@vanlissaWORCSWorkflowOpen2020]

All code and synthetic data available at

<https://github.com/cjvanlissa/veni_forest>

# Results

## Variable importance

## Marginal association


## Marginal association 2


## Marginal association 3


## Marginal association 4


## Summary

Important:

* Best predictors: Personality and related constructs
* Best parental predictor is autonomy support (bal. rel.)
  + Negative/intrusive parenting more predictive than other positive practices
* Conflict frequency and behavior
* Empathy, particularly personal distress

Less important than expected:

* SES
* Bullying/victimization
* Delinquency
* Substance use
* Monitoring

## Reflection content

* Most important predictors are routinely assessed (big 5) or overt (conflict)
  + Prime candidates for early risk assessment
  + Conflict resolution behavior can be taught
    - Target for intervention if association is causal
* Emphasis on parenting (in literature and theory) might not be justified
  + Congruent with RI-CLPM showing few parenting effects in adolescence [@vanlissaRoleFathersMothers2019]
* Order of predictors mirrors bioecological model


## Reflections form

* Many predictors show non-linear effects: Emotion dysregulation only for +/-1SD
* Some predictors show almost no marginal effect (e.g., father's age, drug use)
  + This suggests they might be important in **interactions**

## Limitations

* Selective sample
  + WEIRD, Dutch, mostly mid- high-SES (restriction of range)
* Panel study, so few biological / cognitive / neurological measures
* Relatively small sample size
  + If you have access to 1000+ sample, please get in touch
* Computational constraints prevented us from tuning the model beyond default settings, and from exploring marginal effects for higher order interactions

## Conclusions

* Proximal factors more predictive of ER development than distal predictors
* Parenting may be less relevant than previously thought
* Personality and conflict behavior are candidates for early diagnosis
* Conflict skills training might be avenue for intervention

\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup
