---
title             : "Complementing theory on adolescent emotion regulation using data-driven insights from machine learning"
shorttitle        : "Title"

author:
  - name: "Caspar J. Van Lissa"
    affiliation: "1,2"
    corresponding: yes
    address: "Padualaan 14, 3584CH Utrecht, The Netherlands"
    email: "c.j.vanlissa@uu.nl"
  - name: "Lukas Beinhauer"
    affiliation: "3"
  - name: "Susan Branje"
    affiliation: "4"
  - name: "Wim H. J. Meeus"
    affiliation: "4"

affiliation:
  - id: "1"
    institution: "Utrecht University, department of Methodology & Statistics"
  - id: "2"
    institution: "Open Science Community Utrecht"
  - id: "3"
    institution: "Helmut-Schmidt-Universität, Department of Methodology and Statistics for Psychology"
  - id: "4"
    institution: "Utrecht University, department of Youth and Family"

authornote: |
  Data of the RADAR (Research on Adolescent Development And Relationships) study were used (https://doi.org/10.17026/dans-zrb-v5wp). RADAR has been financially supported by main grants from the Netherlands Organisation for Scientific Research (GB-MAGW 480-03-005, GB-MAGW 480-08-006), Stichting Achmea Slachtoffer en Samenleving (SASS), the Netherlands Organisation for Scientific Research to the Consortium Individual Development (CID; 024.001.003), a grant of the European Research Council (ERC-2017-CoG - 773023 INTRANSITION), and various other grants from the Netherlands Organisation for Scientific Research, VU University Amsterdam, and Utrecht University. The funders had no role in the study design, data collection and analysis, decision to publish, or preparation of the manuscript. If the regulations of the journal allow it, the participating children, parents and schools are to be thanked as well.

abstract: |
  https://www.srcd.org/news/child-development-special-section-formalizing-theories-child-development
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  
  One or two sentences to put the results into a more **general context**.
  
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib", "references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
knit              : worcs::cite_all
---

```{r setup, include = FALSE}
library("papaja")
library("worcs")
# We recommend that you prepare your raw data for analysis in 'prepare_data.R',
# and end that file with either open_data(yourdata), or closed_data(yourdata).
# Then, uncomment the line below to load the original or synthetic data
# (whichever is available), to allow anyone to reproduce your code:
# load_data()
```

<!-- TS: Emotion regulation development is a critical challenge in adolescence. -->
<!-- TS: Emotion dysregulation has severe consequences for individual adolescents, and society at large. -->
<!-- TS: Therefore, substantial research has addressed this topic. -->
<!-- TS: Most preceding research has been limited in scope, focusing on only a few theoretically-driven predictors. -->
<!-- TS: Machine learning makes it possible to cast a wider net, and explore all (available) potentially relevant predictors. -->
<!-- TS: Confirmatory versus exploratory research: Define the distinction. -->
<!-- TS: Some longitudinal research falls between these categories. For example, complex longitudinal studies, where prior research is sufficient to lead us to hypothesize *some* associations between two constructs, but not sufficiently specific to let us hypothesize which parameters will be significant. -->
<!-- TS: Instead, us -->

A key challenge in adolescence is developing mature emotion regulation abilities [@zimmermannEmotionRegulationEarly2014]. 
Adolescents experience biological, cognitive, and social changes that create an opportunity for positive emotion regulation development [@steinbergAgeOpportunityLessons2014].
Still, as many as 20% of adolescents instead experience severe emotion regulation difficulties [@leeAdolescentMentalHealth2014],
with potentially lifelong implications for mental health and well-being [@aldaoEmotionregulationStrategiesPsychopathology2010].
These difficulties in emotion regulation can cause substantial distress to the individual,
and incur a sizable societal cost.
A comprehensive understanding of the risk factors that predispose some adolescents to emotion dysregulation could aid in the development of effective screening and interventions.
However, despite an abundance of empirical research,
many studies are narrow in scope, considering only a few predictors,
the literature is fragmented across subdisciplines, and lacks a unifying theoretical framework
<!-- that is both specific to adolescence and comprehensive in scope  -->
[@vanlissaMappingPhenomenaRelevant2021].

The present study seeks to overcome these limitations
by using data-driven insights to complement and reflect on existing theory.
To this end, a broad range of potential risk factors were analyzed using machine learning,
and ranked in terms of predictive importance.
<!-- These predictors may include both risk factors that render some youth susceptible to emotion regulation difficulties, and correlates that can be used to identify at-risk youth. -->
Specifically,
the SEM-forests algorithm was used to
identify the most important early indicators of developmental trajectories
of emotion regulation across adolescence.
The resulting insights indicate which factors are most predictive of adolescents' diverging destinies in terms of emotion regulation development.
Armed with this knowledge, future research and interventions can focus on the most important predictors.
Moreover, these insights can guide the formation of overarching theory that adequately represents essential phenomena.
<!-- complement existing theory on adolescent emotion regulation using data-driven insights from machine learning. -->

## Emotion dysregulation in adolescence

Emotion regulation,
defined as the ability to manage emotional experience and expression,
is critical for mental health [@aldaoEmotionregulationStrategiesPsychopathology2010, @schaferEmotionRegulationStrategies2017, @braetEmotionRegulationChildren2014]
and social competence [@zimmermannEmotionRegulationEarly2014] throughout life.
There is mounting evidence, however, that adolescents do not uniformly become better at regulating emotions: 
staggered development of emotional and regulatory brain circuits gives rise to a "maturity gap" in mid-adolescence [@croneUnderstandingAdolescencePeriod2012].
The emotional salience of social stimuli is intensified, while the ability to regulate these emotions lags behind [@caseyBrakingAcceleratingAdolescent2011, @vanlissaLongitudinalInterplayAffective2014].
This imbalance motivates most adolescents to engage in adaptive exploration:
Seeking out experiences required to develop mature socio-emotional competences.
At the same time, this imbalance renders some adolescents vulnerable to emotion dysregulation [@leeAdolescentMentalHealth2014],
<!-- exposes youth to a risk of emotion dysregulation, -->
for example, because they are not yet well-equipped to regulate the pursuit of unhealthy incentives,
or persevere in pursuing adaptive goals when facing obstacles.
Thus, some become trapped in a negative spiral characterized by more negative and volatile emotions [@craccoEmotionRegulationChildhood2017],
which has been linked to the emergence of several mental disorders [@schaferEmotionRegulationStrategies2017].
<!-- This imbalance motivates most youngsters to seek out new challenges in life and love, -->
<!-- but renders some of them vulnerable to emotion dysregulation1,10. Unfortunately, there is little theory about which risk factors and environmental hazards render adolescents susceptible to emotional difficulties11. -->
<!-- Acquiring adaptive emotion regulation skills in adolescence plays a key role in this regard7,20, because t -->
 
Empirical support for this notion of diverging destinies is evident from studies that examine developmental trajectories of emotion regulation in adolescence: 
Such studies consistently reveal large between-individual heterogeneity [see @craccoEmotionRegulationChildhood2017; @vanlissaRoleFathersMothers2019; @zimmermannEmotionRegulationEarly2014; @reindlSocializationEmotionRegulation2016].
An important next step is to identify which risk factors render some adolescents susceptible to emotion dysregulation.
<!-- This goal fits within the paradigm shift that has gripped the field of pedagogy, towards “person-centered” research, which focuses on individual developmental trajectories and within-family predictive processes, rather than average patterns across samples16,23. Seminal studies have applied “clustering” methods to identify youth with similar developmental trajectories14,24–26. Adolescents may follow similar trajectories for different reasons, however, and person-centered methods are too complex to accommodate more than a handful of theoretically-derived predictors. Unfortunately, relevant theory on adolescent emotion regulation development is limited27–29. The field is therefore at an impasse: We know that some adolescents are more susceptible to emotion dysregulation than others, but lack tools and theory to identify important predictors of individual development30. -->

## Potential predictors of emotion regulation development

Several theories are commonly cited in research on adolescent emotion regulation.
Notably, there appears to be a trade-off between these theories' scope and specificity.
Some theories are broad in scope, considering many potential risk factors, but are non-specific.
This limits their usefulness in generating testable predictions, and renders them impervious to falsification.
Other theories are narrow in scope, and are sufficiently specific to derive testable predictions, but lack broader context.
Two examples of macro-theories are the bioecological model [@bronfenbrennerBioecologicalModelHuman2007] and the transactional model [@sameroffUnifiedTheoryDevelopment2010].
The bioecological model describes how individual development is shaped by factors at different levels of analysis:
The individual level, which includes biological make-up and individual differences;
the microsystem, composed of primary caregivers and close relationships;
the mesosystem, including school and neighborhood factors;
the exosystem, including media, political- and economic influences;
and the macrosystem, consisting of cultural norms and values.
The transactional model is compatible with the bioecological model;
it describes development as an interplay between the individual child and the environment.
It distinguishes proximal environmental influences, roughly corresponding to the microsystem,
from more distal influences.
With increasing age, distal influences are thought to become more influential. 
More specific theories primarily focus on cognitive maturation and the role of primary caregivers,
including the theory of normative emotional development [@sroufeAttachmentDyadicRegulation1996]
and the tripartite model of emotion regulation socialization by parents [@morrisRoleFamilyContext2007] and peers [@reindlSocializationEmotionRegulation2016].
Note that the tripartite model is one of the few theories that explicitly addresses the life phase of adolescence.
It would be beneficial to work towards a more comprehensive theory of emotion regulation development,
specific to adolescence,
that is broad in scope and still specific enough to derive testable predictions.

<!-- * Bioecological model [@bronfenbrennerBioecologicalModelHuman2007] -->
<!-- * Transactional model [@sameroffUnifiedTheoryDevelopment2010] -->
<!-- * Hall's notion of "storm and stress" -->
<!-- * Theory of normative emotional development [@sroufeEmotionalDevelopmentOrganization1995] -->
<!-- * Tripartite model [@morrisRoleFamilyContext2007] -->
<!-- * Internalization model [@holodynskiDevelopmentEmotionsEmotion2006] -->
<!-- * Polyvagal theory [@porgesOrientingDefensiveWorld1995]. -->
<!-- * Model of social-affective engagement and goal flexibility [@croneUnderstandingAdolescencePeriod2012] -->
<!-- * Process model of emotion regulation [@grossHandbookEmotionRegulation2013] -->
<!-- * Social Information Processing Theory [@lemeriseIntegratedModelEmotion2000] -->

<!-- Maybe review theory briefly? -->
The first step towards theory formation or revision is identifying relevant phenomena [@borsboomTheoryConstructionMethodology2020].
Thus, offering a starting point for this endeavor,<!-- empirical investigations of predictors of adolescents' emotion regulation development -->
a recent literature review 
used a new method, the text mining systematic review, to identify relevant phenomena in relation to adolescents' emotion regulation based on the analysis of author keywords and abstracts of 6305 publications  [@vanlissaMappingPhenomenaRelevant2021].
Typically, phenomena are identified using narrative literature reviews.
However, these are often limited to a small number of papers and structured around well-established ideas [@littellEvidencebasedBiasedQuality2008]
Text mining systematic reviews, by comparison, are relatively comprehensive and derive insights from the literature through a relatively transparent and objective - albeit less interpretative - procedure [@vanlissaMappingPhenomenaRelevant2021]. <!-- reflection and meaning-making-->

The results of that review indicated that internalizing and externalizing problems and mental health-related outcomes were most frequently studied in conjunction with emotion regulation. 
Some relevant phenomena reflected themes also identified in prior narrative reviews,
including cognitive development, parenting, and external stressors [cf. @coe-odessEmergentEmotionsAdolescence2019].
Other relevant phenomena were underrepresented in prior theory and reviews,
such as substance use, structural disadvantage,
identity, morality, sex and sexual development.
This relatively comprehensive overview offers a starting point for empirical investigations of predictors of adolescents' emotion regulation development, such as the present study.
<!-- To this end, the present study sought to identify operationalizations of these relevant phenomena in a large-scale panel study. -->

## Complementing theory using machine learning

According to models of knowledge production within the quantitative positivist paradigm,
a field progresses through the cyclical alternation of exploratory and confirmatory research, also known as the "empirical cycle" [@degrootMethodologieGrondslagenVan1961, @wagenmakersCreativityVerificationCyclePsychological2018].
Exploratory research gives rise to testable predictions,
and confirmatory research evaluates testable predictions on new data.
Crucially, these phases are linked through evaluation of existing theory in light of empirical findings [@degrootMethodologieGrondslagenVan1961].
The fact that our field still lacks overarching theory despite decades of productive empirical research
suggests that the empirical cycle is not routinely completed.
In support of this claim, it has been noted that the overwhelming majority of research in our field appears, at first glance, to be confirmatory [@demuthQualitativeMethodologyDevelopmental2015].
More exploratory research is thus needed to complete the cycle.
Secondly, and paradoxically given the first argument,
the link between confirmatory research and theory is often tenuous.
For instance, many studies derive their predictions from prior empirical research, not from theory [e.g., @vanlissaCostEmpathyParentadolescent2017].
Thirdly, theories are rarely falsified or revised in light of empirical findings.
This failure to evaluate theory leads to flexible, "invulnerable" frameworks that can accommodate any result [@szollosiArrestedTheoryDevelopment2021].
To advance theory formation, our field might benefit from "closing the empirical cycle" through increased exploratory research and reflection on theory.

There are preliminary indications of a move towards exploratory research:
Developmental scientists are turning to increasingly complex structural equation models.
Consider, for example, the burgeoning interest in disentangling between-family differences from within-family processes using the innovative random-intercept cross-lagged panel model [@hamakerCritiqueCrosslaggedPanel2015] and dynamic structural equation modeling [@asparouhovDynamicStructuralEquation2018].
Many authors using these methods have realized that existing theory is not sufficiently precise to derive distinct hypotheses at these two levels [see @nevilleBidirectionalAssociationsScreen2021; @vanlissaMothersFathersQuantitative2020].
By necessity, these models are thus used in at least a partly exploratory capacity.
This approach has two pitfalls.
Firstly, complex models are by definition more flexible and fit data better then simpler models, at the expense of decreased generalizability [@hastieElementsStatisticalLearning2009].
Flexibility is a hallmark of exploratory research [@wagenmakersCreativityVerificationCyclePsychological2018],
but increases the risk of false-positive results and questionable research practices in confirmatory research [@wichertsDegreesFreedomPlanning2016, @simmonsFalsePositivePsychologyUndisclosed2011].
It is therefore essential that flexibility is disclosed, and that exploratory and confirmatory research are clearly demarcated.
Secondly, structural equation models are inherently a confirmatory method,
and they can only include a limited number of predictors due to concerns of statistical power [@klinePrinciplesPracticeStructural2005].
Their use for exploration is limited, effortful, and potentially fraught (e.g., when using p-values for discovery instead of inference).
We argue that machine learning offers a suitable tool for rigorous exploration:
These methods discover patterns in data while maximizing generalizability,
and incorporate checks and balances to curtail false-positive results [@hastieElementsStatisticalLearning2009].
In the context of the present study, we use a specific machine learning algorithm uniquely suited for theory-guided exploration: SEM-forests [@brandmaierTheoryguidedExplorationStructural2016].

<!-- * Predictors of latent growth model? -->
<!--   + Only linear differences, few predictors (power) -->
<!-- * Multigroup latent growth model? -->
<!--   + Non-linear differences, but only one moderator -->
<!-- * Latent class growth model with auxilliary variables? -->
<!--   + Groups with regard to trajectory, not predictors -->

<!-- ## Solution: Machine learning -->

<!-- * Allows non-linear differences between trajectories -->
<!-- * Performs variable selection -->
<!--   + Cast wide net among potentially relevant predictors -->
<!-- * Checks & balances ensure generalizability -->
<!-- * Exploratory: No hypothesis -->


## The present study

The present study set out to identify the most important early indicators of developmental trajectories
of emotion regulation across adolescence.
This involves three important challenges:
First, the number of potentially relevant risk factors, as identified in a prior text mining systematic review, is large.
Second, there is a paucity of relevant theory to narrow down the candidate predictors.
Third, methods commonly used in the field are not well suited for large numbers of predictors.
Machine learning methods that perform variable selection can overcome these challenges.
We therefore used the SEM-forests algorithm to identify the most important early predictors of trajectories of emotion regulation development throughout adolescence.
The primary aim is to identify the most important predictors of adolescents' trajectories of emotion regulation development.
The secondary aim is to probe the nature of the marginal association of the predictors with these trajectories.


<!-- Deductive research, or hypothesis-testing, is the predominant focus of contemporary behavioral research. It tends to focus on a relatively narrow set of theoretically-derived variables, and the results revolve around statistical inference: Whether the theoretical hypotheses are supported by significant or reliable effects. In deductive research, less emphasis is placed on comprehensiveness or breadth of candidate predictors. Relatedly, the relative importance of different predictors is often of secondary importance, as is the model’s predictive performance. Thus, although an advantage of deductive approaches is that they can be used to draw inferences about theoretical hypotheses, they also have specific limitations.  -->
<!-- To allocate scientific resources effectively, it is important to cast a wide net among potential predictors, across different theories, and even include under-theorized factors to unearth potential blind spots in the extant literature. Inductive research – that is, rigorous exploratory work that identifies reliable patterns in data, is more suited to these demands.  -->
<!-- In recent years, inductive research has been gaining traction as a technique to complement existing theories by identifying important omissions[2]. In particular, machine learning offers powerful new tools for systematic exploration that can identify relevant predictors and complex relationships that have eluded theoreticians[5]. Machine learning is an approach to data analysis that focuses on maximizing predictive performance. This involves the use of flexible models to find reliable patterns in data. Machine learning models can distill a large set of candidate variables down to the ones that are most important in predicting the outcome of interest, and also indicate the direction and shape of the marginal association between those predictors and the outcome. In a context where predictor variables are likely to be related to each other, machine learning is better suited to manage these complex relationships than, e.g., multiple regressions. Moreover, it incorporates checks and balances to prevent spurious findings [3]. -->




<!-- The field of adolescent emotion regulation is prolific, but lacks overarching theory. -->
<!-- For guidance on how to remedy this shortcoming, we turn to philosophy of science. -->
<!-- De Groot’s "empirical cycle" constitutes a model of the production of knowledge through scientific research [@degrootMethodologieGrondslagenVan1961]. -->
<!-- It consists of five phases: *observation* of phenomena to be explained; *induction* of hypotheses; *deduction* or derivation of testable predictions; *testing* these predictions in new data; and finally *evaluation*: reflecting back on the theories and hypotheses that gave rise to the tested predictions. -->
<!-- The model is cyclical because *evaluation* offers new interpretative data for the researcher to observe - such as rejected hypotheses and serendipitous findings. -->

<!-- The empirical cycle is merely one model, -->
<!-- and there is a pluriformity of other epistemological approaches to knowledge production. -->
<!-- Nonetheless, u -->
<!-- Unsubstantiated or undisclosed deviations from the empirical cycle may signal practices that limit scientific progress. -->
<!-- Consider three examples: -->
<!-- First, conducting many analyses and then writing a paper that presents significant findings as tests of a-priori hypotheses, or "hypothesizing after the results are known", constitutes a questionable research practice that leads to false-positive results [@wichertsDegreesFreedomPlanning2016]. -->
<!-- Second, failure to engage in evaluation and reflect back on theory in light of empirical findings leads to flexible, "invulnerable" theories that can accommodate any result [@szollosiArrestedTheoryDevelopment2021]. -->
<!-- Third, and particularly relevant to the present study, -->
<!-- is failure to complete the cycle. -->
<!-- In a productive field of study, the empirical cycle is regularly completed; -->
<!-- evaluation of theories in the light of empirical findings allows scholars to hone those theories, -->
<!-- which in turn yield increasingly precise predictions. -->
<!-- We argue that, the paucity of overarching theory in our field can be remedied by completing the empirical cycle. -->

<!-- The phases of the empirical cycle map onto two modes of conducting research: -->
<!-- Confirmatory research, which involves the phases of deduction, testing, and evaluation, and exploratory research, which involves the phases of observation and induction  [@wagenmakersCreativityVerificationCyclePsychological2018]. -->
<!-- Although developmental psychology is dominated by quantitative positivism, -->
<!-- relatively few studies are explicitly exploratory [@demuthQualitativeMethodologyDevelopmental2015]. -->
<!-- There are two main indications that some studies are implicitly exploratory. -->

<!-- The first is that some, apparently confirmatory, papers are not rooted in any theoretical framework, but instead based exclusively on prior empirical literature. -->

<!-- The second indication is that an increasing number of studies use complex modeling techniques in the absence of theory specific enough to derive hypotheses for such models. -->
<!-- For example, consider the recent literature using the random-intercept cross-lagged panel model, -->
<!-- a complex structural equation model that separates stable differences between individuals from processes within  individuals (or families). -->
<!-- Several authors have remarked that theory in their areas of expertise is insufficiently precise to derive specific and distinct predictions about effects at the between- and within level. -->
<!-- At first glance, these publications fit the mold of confirmatory research, providing hypotheses and tests thereof. -->
<!-- Upon further reflection, the hypotheses tested are not very specific (e.g., specifying a positive relation between two constructs, regardless of level of analysis), -->
<!-- and they can be simultaneously supported and refuted by different parameters in the model without repercussions for theory. -->
<!-- They are thus not truly "testable" in the sense that De Groot intended. -->
<!-- Furthermore, the focus in these publications is on post-hoc interpretation of many parameter estimates - a hallmark of exploratory research. -->
<!-- Implicitly exploratory research such as this suggests that others have perceived the need for more exploration, -->
<!-- but that exploratory work is wrt as confirmatory the exploratory mode of research is not ingrained in the field. -->


<!-- have noted that existing theory is insufficiently specific  -->
<!-- ,  -->

<!-- and devoting greater attention to exploratory research and theory formation. -->



<!-- In this model, the hypothesis plays an important role. -->
<!-- It is a testable prediction that can be supported or falsified; depending on the results, we reflect on old knowledge – which includes theory – and make necessary adjustments that lead to new hypotheses. -->
<!-- A hypothesis can thus be a specific prediction that delivers new knowledge when it is false; -->
<!-- this drives a scientific field forward. -->


<!-- Prior theory and empirical research thus suggests many phenomena potentially related to heterogeneity in adolescents' emotion regulation development. -->
<!-- What is missing is an understanding of the relative utility of each of these phenomena in predicting developmental outcomes. -->
<!-- Individual studies do not suffice for this purpose as -->
<!-- they tend to sample from different populations, -->
<!-- use different research designs, -->
<!-- different control variables and statistical models. -->
<!-- Relatedly, many studies tend to examine only a handful of predictors, -->
<!-- often guided by a theory with narrow scope -  -->
<!-- , it is hard to tell how important the variables are, relative to other variables considered by other theories (or variables not considered at all). Machine learning is a more holistic methodology, as it can assess and compare a large number of potential predictors simultaneously, including theoretically relevant ones, and identify which predictors ultimately explain the most variance in the outcome measure of interest -->


<!--   Machine learning is a more holistic methodology, as it can assess and compare a large number of potential predictors simultaneously, including theoretically relevant ones, and identify which predictors ultimately explain the most variance in the outcome measure of interest. -->


<!-- Theory cons methods for theory construction state that the first step in this endeavor is to identify well-established relevant *empirical phenomena*, -->
<!-- defined as stable and general features of the world [@borsboomTheoryConstructionMethodology2020]. -->
<!-- The current study used a text mining systematic review to provide a comprehensive map of the empirical phenomena relevant to adolescent emotion regulation. -->

<!-- Currently, few instruments exist for the identification of relevant phenomena; instead, theorists typically rely on expert opinion [@borsboomTheoryConstructionMethodology2020]. -->
<!-- This paper argues that systematic reviews are a suitable and relatively more objective method for phenomena detection. -->
<!-- Traditional narrative reviews are limited, however, by small convenience samples, -->
<!-- confirmation bias, -->
<!-- and an undue emphasis on positive results [@littellEvidencebasedBiasedQuality2008]. -->
<!-- These limitations are overcome by a novel method: -->
<!-- the text mining systematic review (TMSR). -->
<!-- Relative to narrative reviews, text mining can cover arbitrarily greater corpora, -->
<!-- and gleans insights from the literature by means of a relatively more transparent, objective, and reproducible process. -->
<!-- <!--Where sentient readers tend to structure their reading of the literature around established ideas [@littellEvidencebasedBiasedQuality2008], -->
<!-- the inherent "fairness" of text mining gives emerging themes a chance to come to the fore.--> -->
<!-- This approach assumes the frequency with which phenomena recur in the literature to be indicative of their relevance. -->
<!-- If it is additionally assumed that the frequency with which phenomena are studied together is indicative of the extent to which they are thought to be associated, -->
<!-- then the co-occurrence of phenomena within publications can be interpreted as a rudimentary nomological network: a mapping of relationships between theoretically relevant phenomena [see @alaviAligningTheoryMethodology2018]. -->
<!-- Such a network can serve as a starting point for the formulation of a proto-theory, -->
<!-- which would additionally involve abductive reasoning: specifying putative mechanisms to explain associations between the phenomena [@borsboomTheoryConstructionMethodology2020]. -->
<!-- Text mining systematic reviews constitute an inductive and exploratory approach: -->
<!-- Their aim is phenomena detection based on generalization from observations (data) without appeal to theory [@haigExploratoryFactorAnalysis2005]. -->
<!-- By identifying and mapping the phenomena relevant to adolescent emotion regulation, -->
<!-- such an inductive approach lays the groundwork for future theory development. -->

<!--
stifterEmotionRegulation2019
* call for studies incorporating predictors of ER across multiple levels of analysis
* different conceptual frameworks

riedigerEmotionRegulationAdolescence2014
* difficulty in deriving a cohesive picture about emotion regulation in adolescence
* different conceptual frameworks

* three themes (neurocognitive development, socialization, interplay between internal and external factors) currently represent relatively independent bodies of research
* A more explicit integration of these various research perspectives will help us to arrive at a more integrated and interpretable picture of the development of emorion regulation during adolescence. 

@bussTheoriesEmotionalDevelopment2019

* emotional development theory is still in its own state of development. 
* empirical evidence can contribute to theory development.

@bariolaChildAdolescentEmotion2011
* Further research apply reﬁned theoretical conceptualizations of ER.


Where sentient readers tend to structure their reading of the literature around established ideas [@littellEvidencebasedBiasedQuality2008],
This potentially reveals phenomena and associations missed by narrative reviewers,
thereby laying the groundwork for future empirical and theoretical work.


we set out to map these relevant empirical phenomena using a systematic review.


generate a proto-theoretical nomological network of the relevant empirical phenomena.
We discuss how this nomological network relates to existing relevant theory, and conclude
with recommendations for future theory-generating efforts.
There have been calls for the  consolidation of these disparate insights into overarching theory would be an important step forward.
An important first step 
and intrinsic limitations of narrative reviews prohibit comprehensive coverage of the vast but diffuse literature .-->
<!--
Recent years have seen the publication of several excellent systematic reviews,
but these have all been narrative reviews.-->

 <!-- CJ To fill out gaps, following up on promising constructs and unify heretofore disparate regions of the lit -->
<!-- SH: this implies that the main contribution of the paper is predominantly methodological is that the case? Is there some primary thesis/argument that will be advanced? It is really important to take a top-down approach to that messaging, beginning here in the first paragraph-->
<!-- Ergens anders: Similarly, although
there are several *relevant* theories, there is no specific theoretical
framework to guide research on adolescents' emotion regulation. Although theory
development is beyond the scope of this paper, we set out to generate a
proto-theoretical nomological network of the potential risk factors,
manifestations, and outcomes of emotion regulation in adolescence. We discuss
how this nomological network relates to existing relevant theory, and conclude
with recommendations for future theory-generating efforts.-->
<!-- At the same time, the staggered development of motivational-emotional and regulatory brain circuits gives rise
to a maturity gap [@croneUnderstandingAdolescencePeriod2012]. 
This leads youngsters to pursue new experiences in life and love,
without being fully prepared to cope with the emotional outcomes. --> 

<!--The literature thus paints a picture of
adolescence as a chrysalis for emotional development: Children enter this stage
with emotion regulation skills adapted to the challenges of childhood. During
adolescence, emotional systems are rearranged substantially.-->

<!--
* intense negative emotions in daily life (Silk, Steinberg, & Morris, 2003) 
* unstable peer or romantic relationships (Furman & Collins, 2009)
* decrease in perceived support from parents (Furman & Buhrmester, 1992).
* early adolescence is characterized by a higher rate of conflicts with parents (Laursen, Coy, & Collins, 1998)
* higher variability of negative emotions compared to late adolescence (Larson, Moneta, Richards, & Wilson, 2002).
* relative to other developmental stages, adolescence involves a particularly large number of transitions, novel situations, and new stimuli, physical, cognitive, emotional, and social development—particularly in early and mid-adolescence (Seiffge-Krenke, 2000)
-->

<!-- ## Knowledge gap -->

<!-- Substantial empirical research, but: -->

<!-- * Different (sub)disciplines have approached the topic in disparate ways [@riedigerEmotionRegulationAdolescence2014] -->
<!-- * No consistent terminology [@bariolaChildAdolescentEmotion2011] -->
<!-- * No conceptual frameworks [@stifterEmotionRegulation2019] -->
<!-- * No overarching theoretical framework [@bussTheoriesEmotionalDevelopment2019] -->

<!-- Knowledge must be consolidated into overarching theory -->


## Method


SEM-forest [@brandmaierTheoryguidedExplorationStructural2016]

1. Bootstrap sample
2. On each sample, estimate a SEM-model
3. Consider $k$ candidate predictors
4. Identify predictor and value that maximizes $LR$ of post-split multi-group model
5. Average predictions across bootstrap samples

We used 1000 bootstrap samples, $k = 9$


**Participants**

* RADAR data (https://www.uu.nl/en/research/radar)
* 497 Dutch adolescents (283 boys; age at T1: M =13.03, SD = 0.46)
* Collected between 2006-2011
* Most families were  medium- to high-SES (10% low-SES)

## Candidate predictors

Predictors: 87

Categories: 

* Demographics (e.g., SES)
* Biological (age, sex, pubertal development)
* Individual differences (big5, BIS/BAS, empathy)
* Risk behavior (in/externalizing symptoms, substance use, delinquency, school performance, victimization)
* Relationship quality (parents/peers)
* Parenting (perceived and parent-reported)
* Conflict resolution styles

## Dependent variable

Difficulties in emotion regulation [@gratzMultidimensionalAssessmentEmotion2004]

* 24 items (omitted 'emotional awareness' subscale)
* E.g.: “When I’m upset, I have difficulty thinking about anything else”
* 1 (“Almost never”) to 5 (“Almost always”)
* Excellent reliability, $(.94 > \alpha > .95)$.
* Metric invariance across waves
* Quadratic growth model had good fit
  + CFI = .996, TLI = .993, RMSEA = .047, SRMR = 0.020

## Open science

Workflow for Open Reproducible Code in Science (WORCS) used to make analyses reproducible [@vanlissaWORCSWorkflowOpen2020]

All code and synthetic data available at

<https://github.com/cjvanlissa/veni_forest>

# Results

## Variable importance

## Marginal association


## Marginal association 2


## Marginal association 3


## Marginal association 4


## Summary

Important:

* Best predictors: Personality and related constructs
* Best parental predictor is autonomy support (bal. rel.)
  + Negative/intrusive parenting more predictive than other positive practices
* Conflict frequency and behavior
* Empathy, particularly personal distress

Less important than expected:

* SES
* Bullying/victimization
* Delinquency
* Substance use
* Monitoring

## Reflection content

* Most important predictors are routinely assessed (big 5) or overt (conflict)
  + Prime candidates for early risk assessment
  + Conflict resolution behavior can be taught
    - Target for intervention if association is causal
* Emphasis on parenting (in literature and theory) might not be justified
  + Congruent with RI-CLPM showing few parenting effects in adolescence [@vanlissaRoleFathersMothers2019]
* Order of predictors mirrors bioecological model


## Reflections form

* Many predictors show non-linear effects: Emotion dysregulation only for +/-1SD
* Some predictors show almost no marginal effect (e.g., father's age, drug use)
  + This suggests they might be important in **interactions**

## Limitations

* Selective sample
  + WEIRD, Dutch, mostly mid- high-SES (restriction of range)
* Panel study, so few biological / cognitive / neurological measures
* Relatively small sample size
  + If you have access to 1000+ sample, please get in touch
* Computational constraints prevented us from tuning the model beyond default settings, and from exploring marginal effects for higher order interactions

## Conclusions

* Proximal factors more predictive of ER development than distal predictors
* Parenting may be less relevant than previously thought
* Personality and conflict behavior are candidates for early diagnosis
* Conflict skills training might be avenue for intervention

\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup
